---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
library(knitr)
```

### Load data

```{r load-data}
load("movies.Rdata")
```
* * *

## Part 1: Data
The data set is comprised of 651 randomly sampled movies produced and released before 2016.
As this data is observational, not experimental, there are chances of collinearity in the Multiple Linear Regression Model.
* * *

## Part 2: Research question

The research question is:
What are the attributes that make movie popular?

This question is of interest because it can help predict the success of the movie based on it's attributes.

* * *

## Part 3: Exploratory data analysis
For EDA purpose, we are eliminating some variables, which are not relevant and keeping only the following ones(15) to take forward our analysis.
genre, runtime,mpaa_rating,critics_score,critics_rating,audience_score,audience_rating,imbd_rating,imdb_num_votes,best_pic_nom,best_pic_win,best_dir_win,best_actor_win,best_actress_win,top__box

For this purpose, we have created a new variable "Sub_movies"
```{r}
Sub_movies<-movies%>%select(-c(director:rt_url,thtr_rel_year:dvd_rel_day,studio,title,title_type))
Sub_movies[1:5,1:5]
```

The above gives a sneak peak into the new data frame.

Now, for determining popularity of the movie we are taking combination of two variables "imdb_rating" and "imdb_num_votes"
We create a new variable "movie_popularity" which is created by multiplying "imdb_rating" and log of "imdb_num_votes"

We take log of "imdb_num_votes" here because it will give a better explaination of the combination rather than simply multiplying, also it makes the response variable smaller without taking away it's properties.
We add the new variable to "Sub_movies"

```{r}
Sub_movies$movie_popularity<-movies$imdb_rating*log(movies$imdb_num_votes)
```
###Exploratory Data Analysis
Now, Let's plot some graphs!

```{r}
ar<-ggplot(data=Sub_movies, aes(x=audience_score ,y=movie_popularity, colour=audience_rating))+geom_point()
cr<-ggplot(data=Sub_movies, aes(x=critics_score ,y=movie_popularity, colour=critics_rating))+geom_point()
rt<-ggplot(data=Sub_movies, aes(x=runtime ,y=movie_popularity,colour=critics_rating))+geom_point()

amp<-ggplot(data=Sub_movies, aes(x=audience_score ,y=movie_popularity, colour=mpaa_rating))+geom_point()

mpr<-Sub_movies%>%group_by(mpaa_rating)%>%ggplot(aes(x=mpaa_rating,y=movie_popularity))+geom_boxplot()
g<-Sub_movies%>%group_by(genre)%>%ggplot(aes(x=genre,y=movie_popularity))+theme(axis.text.x = element_text(angle = 90, hjust = 1))+geom_boxplot()
pn<- ggplot(data=Sub_movies, aes(x=best_pic_nom ,y=movie_popularity))+geom_boxplot()
pw<- ggplot(data=Sub_movies, aes(x=best_pic_win ,y=movie_popularity))+geom_boxplot()
amw<-ggplot(data=Sub_movies, aes(x=best_actor_win ,y=movie_popularity))+geom_boxplot()
afw<-ggplot(data=Sub_movies, aes(x=best_actress_win ,y=movie_popularity))+geom_boxplot()
dw<-ggplot(data=Sub_movies, aes(x=best_dir_win ,y=movie_popularity))+geom_boxplot()
bo<-ggplot(data=Sub_movies, aes(x=top200_box ,y=movie_popularity))+geom_boxplot()

par(mfrow=c(4,3))
grid.arrange(ar,cr,rt,amp,mpr,pn,pw,amw,afw,dw,bo)
par(mfrow=c(1,1))
g
```

1.From the above graphs we see there is a clear relationship in popularity of the movie and   variables like audience_rating,critics_rating,best_pic_nom,best_pic_win,top200_box

2.But the relationsip between popularity and variables like mpaa_rating,genre,runtime,best_actor_win,best_actress_win,best_dir_win  is a little vague.

3.We can get better idea of these vague relationships by using some summary statistics on  them.

###Summary Statistics
```{r}
genre_summary <-Sub_movies%>%group_by(genre) %>%summarise(movie_popularity_median=median(movie_popularity), num_median=median(imdb_num_votes),imdb_rating_median=median(imdb_rating),critic_score_median=median(critics_score),audience_score_median=median(audience_score),total_count=n()) %>%arrange(desc(movie_popularity_median))
mpaa_rating_summary <-Sub_movies%>%group_by(mpaa_rating) %>%summarise(movie_popularity_median=median(movie_popularity), num_median=median(imdb_num_votes),imdb_rating_median=median(imdb_rating),critic_score_median=median(critics_score),audience_score_median=median(audience_score),total_count=n()) %>%arrange(desc(movie_popularity_median))
bamw_summary <-Sub_movies%>%group_by(best_actor_win) %>%summarise(movie_popularity_median=median(movie_popularity), num_median=median(imdb_num_votes),imdb_rating_median=median(imdb_rating),critic_score_median=median(critics_score),audience_score_median=median(audience_score),total_count=n()) %>%arrange(desc(movie_popularity_median))
bafw_summary <-Sub_movies%>%group_by(best_actress_win) %>%summarise(movie_popularity_median=median(movie_popularity), num_median=median(imdb_num_votes),imdb_rating_median=median(imdb_rating),critic_score_median=median(critics_score),audience_score_median=median(audience_score),total_count=n()) %>%arrange(desc(movie_popularity_median))
bdw_summary <-Sub_movies%>%group_by(best_dir_win) %>%summarise(movie_popularity_median=median(movie_popularity), num_median=median(imdb_num_votes),imdb_rating_median=median(imdb_rating),critic_score_median=median(critics_score),audience_score_median=median(audience_score),total_count=n()) %>%arrange(desc(movie_popularity_median))
genre_summary
mpaa_rating_summary
bamw_summary
bafw_summary
bdw_summary
```
The above summary statistics have been run on the median of the variables and arranged on the basis of decreasing order of median movie popularity

1. In the first table, we see avg score of movie with genre 'Animation' is highest but highest imdb_rating are scored by genre 'Musical & Performing Arts' and 'Documentary'
This may be because the former has more number of votes than the latter two.

2.Second table shows movies with 'G' certification have scored higher on avg. but imdb_rating of 'Unrated' is highest. But, as seen previously, 'Unrated' has significant lower votes than 'G'

3.The third, fourth and fifth table signifies there is a slight difference(increament) in movie popularity and imdb_rating caused if actor/actress wins award but there is significant difference(increament) in popularity if directors wins

```{r}
top200<-Sub_movies%>% filter(top200_box=="yes")
table(top200$genre)
```



* * *

## Part 4: Modeling
*For modeling purpose we are using:
genre, runtime,mpaa_rating,critics_score,audience_score,best_pic_nom,best_pic_win,best_dir_win,best_actor_win,best_actress_win,top__box

*audience_rating and critics_ratings are excluded because of possible collinearity.

*Although audience_score and critcs_score, best_pic_nom and best_pic_win gives us a hint of collinearity as well, but we will still include them and create a model excluding them later.

We are using "Backward Elimination" method for constructing the model using "Adjusted R squared" as criteria for elimination purpose.

< 1 >
For the initial step, let's take the full model and eliminate variable one by one with replacement.
After it we make a data frame indicating adj r squared values when we eliminate particular variable
```{r}
mlr_model0<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model1<- lm(movie_popularity~runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model2<- lm(movie_popularity~genre+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model3<- lm(movie_popularity~genre+runtime+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model4<- lm(movie_popularity~genre+runtime+mpaa_rating+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model5<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model6<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_win+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model7<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_actor_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model8<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actress_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model9<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model10<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+top200_box,data=Sub_movies)
mlr_model11<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_actress_win+best_dir_win,data=Sub_movies)

adr_df1<-data.frame(eliminated=c("None", "genre"," runtime", "mpaa_rating", "critics_score", "audience_score", "best_pic_nom", "best_pic_win", "best_actor_win",  "best_actress_win","best_dir_win", "top200_box"), Adjusted_RSquared=c(summary(mlr_model0)$adj.r.squared, summary(mlr_model1)$adj.r.squared, summary(mlr_model2)$adj.r.squared, summary(mlr_model3)$adj.r.squared, summary(mlr_model4)$adj.r.squared, summary(mlr_model5)$adj.r.squared, summary(mlr_model6)$adj.r.squared, summary(mlr_model7)$adj.r.squared, summary(mlr_model8)$adj.r.squared, summary(mlr_model9)$adj.r.squared, summary(mlr_model10)$adj.r.squared,summary(mlr_model11)$adj.r.squared) )
adr_df1
```
We observe variable "best_actress_win" is bringing down Adjusted R Squared value.So, we eliminate it and again build our model 

<2>

```{r}
mlr_model0<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model1<- lm(movie_popularity~runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model2<- lm(movie_popularity~genre+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model3<- lm(movie_popularity~genre+runtime+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model4<- lm(movie_popularity~genre+runtime+mpaa_rating+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model5<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model6<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model7<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model8<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_dir_win+top200_box,data=Sub_movies)
mlr_model9<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+top200_box,data=Sub_movies)
mlr_model10<- lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+audience_score+best_pic_nom+best_pic_win+best_actor_win+best_dir_win,data=Sub_movies)

adr_df2<-data.frame(eliminated=c("None", "genre"," runtime", "mpaa_rating", "critics_score", "audience_score", "best_pic_nom", "best_pic_win", "best_actor_win","best_dir_win", "top200_box"), Adjusted_RSquared=c(summary(mlr_model0)$adj.r.squared, summary(mlr_model1)$adj.r.squared, summary(mlr_model2)$adj.r.squared, summary(mlr_model3)$adj.r.squared, summary(mlr_model4)$adj.r.squared, summary(mlr_model5)$adj.r.squared, summary(mlr_model6)$adj.r.squared, summary(mlr_model7)$adj.r.squared, summary(mlr_model8)$adj.r.squared, summary(mlr_model9)$adj.r.squared, summary(mlr_model10)$adj.r.squared) )
adr_df2
```

Now , we see we are getting best value of adjusted r squared without elimination of any variable. Although, 'best_acor_win' is quite close, but we would not eliminate it.

So, 'mlr_model0' is our final model
Let's look into it.
```{r}
summary(mlr_model0)
summary(mlr_model0)$adj.r.squared
```
So, we get Adjusted r Squared value of about 0.623

Now, as we talked about other possible case of collinearity in:
audience_score and critcs_score, best_pic_nom and best_pic_win 

Let's just check the model after eliminating best_pic_nom and audience_score
```{r}

mlr_model_x1<-lm(movie_popularity~genre+runtime+mpaa_rating+critics_score+best_pic_win+best_actor_win+best_dir_win+top200_box,data=Sub_movies)
summary(mlr_model_x1)
summary(mlr_model_x1)$adj.r.squared
```

From this, we ger Adjsuted R Squared to be 0.431, which is a significant drop!

Nevertheless, we will use "mlr_model0"(parsimonious model) for our prediction purposes,which yeilded Adjsuted R Squared to be 0.623

But, before that, let us check whether our model satisfies Multiple Linear Regression conditions.
There are four checks for it.

1.Independence of residuals
  As data was randomly sampled, this condition is met.
  
2. Linearity
```{r}
par(mfrow=c(1,3))
plot(mlr_model0$residuals ~ Sub_movies$audience_score[1:650],xlab="Residuals",ylab="Audience Score")
abline(lm(mlr_model0$residuals ~ Sub_movies$audience_score[1:650]))
plot(mlr_model0$residuals ~ Sub_movies$critics_score[1:650],xlab="Residuals",ylab="Critics Score")
abline(lm(mlr_model0$residuals ~ Sub_movies$critics_score[1:650]))
plot(mlr_model0$residuals ~ Sub_movies$runtime[1:650],xlab="Residuals",ylab="Runtime")
abline(lm(mlr_model0$residuals ~ Sub_movies$runtime[1:650]))
```

From the graphs, we can see, model is linear.

3.Constant Variability of Residuals

```{r}
par(mfrow=c(1,2))
plot(mlr_model0$residuals ~ mlr_model0$fitted.values,xlab="Fitted Values",ylab="Residuals",main="fitted vs residuals")
plot(abs(mlr_model0$residuals) ~ mlr_model0$fitted.values,xlab="Absolute Fitted Values",ylab="Residuals",main="abs fitted vs residuals")
```

Here, we see this condition is also valid, as we see no fan-shape distribution.

4.Nearly normal residuals
```{r}
par(mfrow=c(1,2))
hist(mlr_model0$residuals,xlab="Residuals",main="Residuals Histogram")
qqnorm(mlr_model0$residuals)
qqline(mlr_model0$residuals)
```
There is a very slight skew, but otherwise model seems to be fairly normal.
* * *

## Part 5: Prediction

For prediction purpose, we have taken 2016 blockbuster movie "Deadpool" by Marvel Studios
The movie was of genre Action, Adventure and Comedy. So, we have put it into 'Action & Adventure' genre.
We put the data collected for "Deadpool", in a new variable "movie_2016".
```{r}
movie_2016<-data.frame(critics_score=83,audience_score=90,best_pic_nom="no", best_pic_win="no",best_dir_win="no",best_actor_win="no", genre="Action & Adventure",mpaa_rating="R",runtime=c(108),top200_box="yes")
```
Now, let us predict the interval for popularity score using our model and compare it with original expected value.
```{r}
predict(mlr_model0, newdata=movie_2016, interval="prediction")
original_value<-8*log(744347)
original_value
```
We see, the predicted value is within the predicted interval, slightly towards higher side. Than may because of the pre-release hype of the movie as it was waited by Marvel fans for a long time.
Also, 'Comedy' genre might also have affected it.

References:Rotten Tomatoes- https://www.rottentomatoes.com/m/deadpool  
           IMDB           - https://www.imdb.com/title/tt1431045/?ref_=nv_sr_2
* * *

## Part 6: Conclusion

We conclude following from our study:

1.We see Oscar nomination or wins for the movie contributes heavily to it's popularity. Also, if the movie is in TOP 200 Box office collection list, it is a significant contributer.

2. Animation movies are watched by most of the people and they also score the highest in terms of popularity despite being rated lower than many other genres on imdb.

3.'Drama' and 'Action & Adventure' movies have made it more to TOP 200 BOX Office movies,as compared to 0 times of 'Animation' which is surprising.